Modelling the Trojan Asteroids

### 2020-04-26 Report/draft1.tex
cleaned perturbations_3D ready to submit
updated report clean main function files

%\author{
	%\authorstyle{Christopher Gallagher}}
\author{\authorstyle{Christopher Gallagher} %Candidate 6952R
	\institution{University of Cambridge}}
% Example of a one line author/institution relationship
%\author{\newauthor{John Marston} \newinstitution{Universidad Nacional Autónoma de México, Mexico City, Mexico}}

\date{\today} % Add a date here if you would like one to appear underneath the title block, use \today for the current date, leave empty for no date
\usepackage[english]{babel}
\usepackage{tabularx}
%\usepackage[backend=biber,doi=false]{biblatex}
\addbibresource{references.bib}
%\AtBeginBibliography{\small}
%----------------------------------------------------------------------------------------
\begin{document}

@@

\lettrineabstract{Trojan asteroids are v important. I'm going to make that paragraph a little longer to fill out the space in case this is forming some kind of error. i really hope this is long enough.}

@@

%Use https://leancrew.com/all-this/2016/08/lagrange-points-redux/ to create your own contour plot?
\section{Introduction}
The Jupiter trojans, commonly known as the Trojan asteroids, are two large groups of asteroids that share the planet Jupiter's orbit around the Sun in a 1:1 orbit resonance. These two groups are called the Greeks and the Trojans, named after opposing sides in the mythological Trojan war, and lead/trail Jupiter respectively in its orbit. They correspond to Jupiter's two stable Lagrange points: L\subscript{4}, lying 60° ahead of the planet in its orbit, and L\subscript{5}, 60° behind, with asteroids distributed in two elongated, curved regions around these Lagrangian points.

The first Jupiter trojan, 588 Achilles, was discovered in 1906 by the German astronomer Max Wolf \cite{Nicholson1961}, and a total of 7642 Jupiter trojans have been found as of February 2020 \cite{IAU2020}.

Research into Jupiter's trojan asteroids continues, with the particular focus on their origins reliant on an understanding of their orbit stability \cite{DiSisto2019}, \cite{Nesvorn2018}. This informs studies into their composition \cite{Brown2016}, as travel to these asteroids is considered for their potential in mineral mining \cite{Okada2017} \cite{Levison2016}.

The purpose of this report is to use numerical simulation techniques to investigate the stability of orbits about these Lagrange points, demonstrating the asteroid oscillate about these points under small perturbations and quantifying the absolute distance of the asteroids from the Lagrange point (the wander) during their orbits. The impact of variation in planetary/solar mass on asteroid orbit stability will also be considered. \textit{Signpost what is included in each section?}

@@

\section{Theoretical Background?}

\subsection{Lagrange Points} \label{lagrange}
The asteroids exist at/near Lagrange points, defined in Lagrange's initial analysis of the three-body problem in 1772 \cite{Lagrange1772}, were he demonstrated the existence of five equilibrium points for an object of negligible mass orbiting under the gravitational effect of two larger masses. Three of these equilibrium points, L\subscript{1}-L\subscript{3} lie on the line joining the two masses, while each of the remaining two points, L\subscript{4} and L\subscript{5}, lie at the apex of an equilateral triangle with base equal to the separation of the two masses (see Figure \ref{fig:lagrangepoints}). Despite all these points being potential maxima, stable motion is possible around L\subscript{4} and L\subscript{5} due to the Coriolis force \cite{Lissauer2014}.
%lissauer also says stability is provided that the most massive body has at least 25 times the mass of the secondary

\begin{figure}[h]
	\centering
	\includegraphics[width=\linewidth]{Figures/lagrange_points}
	\caption{The location of the five Lagrange equilibrium points in the circular-restricted three-body problem. The solar and planetary masses are denoted by the large and small filled circles, and the letters P, H, and T denote passing, horseshoe, and tadpole orbits respectively. Note that the two masses form an equilateral triangle with each of the L\subscript{4} and L\subscript{5} points. Reproduced from Marzari et al. \cite{Marzari2002}}
	\label{fig:lagrangepoints}
\end{figure}

Figure \ref{fig:lagrangepoints} depicts orbits about L\subscript{4} and L\subscript{5} (known as tadpole orbits), as well as orbits between Lagrange points are also possible (such as horseshoe orbits), described by Murray et al. \cite{Murray1999}.

This report will focus on tadpole orbits, a well documented feature of Trojan orbits \cite{Garfinkel1983}, \cite{Dermott1981}, and the restricted three-body problem in general. Their distinctive shape is result of a long-period motion about the equilibrium point combining with a short-period oscillation due to Keplerian motion of the asteroid.

Szebehely et al. \cite{Szebehely1969} predict this short period tends to the planetary period in the small planetary mass limit, while the long period is given by

\begin{equation}
T_{long} = T_{P} \sqrt{\frac{4}{27 \mu_{2}}},  
\end{equation}

@@

\subsection{Theoretical Model} \label{theory}
The three body problem, where the dynamics of three interacting bodies are determined given their initial positions and velocities, has no analytical (closed-form) solution in the general case \cite{Barrow2008}.

@@

The system of differential coordinates determine the position and velocity of the asteroids, with two equations per spatial coordinate.
\begin{equation}
\frac{dr_{i}}{dt} = v_{i}, \frac{dv_{i}}{dt} = g_{i}, \quad i = x,y 
\end{equation}

@@

We may also consider a frame rotating at the same speed as the massive bodies. As there is 1:1 orbital resonance between Jupiter and the asteroids, all three bodies are stationary in this frame. This significantly increases the accuracy of numerical simulations, as the exact solution is stationary with no explicit time dependence, rather then requiring an infinite power series \cite{Guglielmi2001}, \cite{LeVeque2007}.

@@

\subsection{Symmetry}
This problem contains a number of symmetries, which could be employed to simplify the problem and reduce the computational load. The Trojan and Greek asteroids are in equivalent positions, so experience the same forces, and the system is rotationally and inversion symmetric, so the choice of initial point and orbit direction is arbitrary. Therefore only the Greeks, orbiting counter-clockwise with perturbations applied at t = 0, need to be investigated.

\subsection{Orbit Geometry}
As the three bodies considered here form an equilateral triangle in the initial equilibrium state, as depicted in Figure \ref{fig:geometry}, we can derive the polar coordinates of each body with respect to the centre of mass about which the bodies orbit.

\begin{figure}
	\centering
	\resizebox{\columnwidth}{!}{%
		\begin{tikzpicture}
		\draw (0,0) -- (6,0) -- (3, {6*sin(60)} ) node[above]{Asteroids} node[midway,right]{$a$} -- cycle node[midway,left]{$a$};
		\draw  (1,0) arc (0:60:1cm)  ;
		%\node [label={[shift = {(10mm,5mm)}]{$\frac{\pi}{3}$}}] {};
		\node [label={[shift = {(13:11mm)}]{$\frac{\pi}{3}$}}] {};
		\node [label={[shift = {(-7mm,-7mm)}]{Sun}}] {};
		\node [label={[shift = {(67mm,-7mm)}]{Planet}}] {};

@@

		\end{tikzpicture}
	}	
	\caption{A geometric depiction of the three-body system, in the case where the planet has a mass equal to a third of the sun, and asteroids are considered at the L\subscript{4} point. $ R_{s} $ and $ R_{p}$ denote the (fixed) radii from the centre of mass (the grey point) to the Sun and the planet respectively, while $ r_{a} $ denotes the radius of the asteroids.}
	\label{fig:geometry}
\end{figure}

@@

\section{Methodology}
This system of coupled first-order ordinary differential equations (ODEs) was solved using the scipy \textit{solve\_ivp} function. The time span was taken as 100 orbits (with 100 points sampled per orbit) unless otherwise stated; this corresponds to 1185 Earth years. Rescaled solar system units are used for mathematical ease, so distances are measured in astronomical units (AU), time in earth years and mass in multiples of the solar mass, and to prevent floating point overflows due to the magnitude of the quantities considered in SI units.

@@

\subsection{Integration Method}
%Put this with theory as "numerical methods"?
%Then all theory could be in methodology?
Within \textit{solve\_ivp}, the default solver is RK45 (an explicit Runge-Kutta method of order 5(4) \cite{Dormand1980}) and gives a deviation in asteroid position (from the Lagrange point) in the order of $ 10^{-4}$ AU in the rotating frame over 50 years. This is larger than expected, indicating the system of equations requires an unreasonably small step size for numerical stability with respect to this method, even in regions where the solution curve is smooth \cite{Lambert1991}. This suggests the system is stiff, and solvers designed for this typically do more work per step, allowing them to take much larger steps, and have improved numerical stability compared to the non-stiff solvers such as RK45 \cite{Byrne1987}.

Instead the stiff "Radau" solver (an implicit Runge-Kutta method of the Radau IIA family of order 5 \cite{Hairer2010}) is used for increased stability \cite{Frank1985}, and achieves a deviation in asteroid position in the order of $ 10^{-13}$ AU instead. This also ensures stability in the stationary frame, with deviations of 0.76\% in asteroid separation from Jupiter over $ 10^{3} $ years, compared to 53\% for the best non-stiff solvers.

@@

\subsection{Programme Structure}
Global constants such as solar mass, and sun\textendash planet separation, along with derived values from these such as orbital period and solar radius from the CoM, are given in an importable python module "\textit{constants}".

Functions to evaluate these coupled differential equation systems are defined in module "\textit{orbits}", while additional functions to evaluate the wander during the orbit (under different sampling routines) are implemented in "\textit{wander}". Further files then import these modules and produce the plots given in this report, fully detailed in appendix \ref{structure}.

To consider a varying planet mass, it was considered preferable to avoid reconstructing all functions to take this as an argument as this requires re-evaluating all initial derived constants. Therefore, I iterate over alternative masses, re-defining constant values in this instance, and then directly import the required functions to compute the orbit. \textbf{REPHRASE?}

Complete code listings are given in appendix \ref{Code}.

\subsection{Performance} \label{Performance}
Sampling 100 points per orbit for 100 orbits takes a mean time of $16.97 \pm 34$ \si{\milli\second}, with sub-linear scaling for sampling rate and orbit number up to the array memory limit, achieved through the optimised integration routines within \textit{solve\_ivp}.

\subsection{Fourier Analysis} \label{Fourier}
For periodic oscillations/orbits, the Fast Fourier Transform (FFT) implemented in \textit{scipy} is used to obtain values for period, with peak finding in \textit{scipy.signal} identifying the exact values. The errors estimated using the associated peak widths of each frequency component.

\section{Results}
\subsection{Unperturbed Stability of Lagrange Points} \label{unperturbed}
Considering the Greeks' orbit without perturbations applied, it has a maximum deviation of $4.68 \times 10^{-13}$ AU over 100 orbits (1185 years). This value is unchanged if 1000 orbits are considered instead, confirming the stability of this Lagrange point.

In the stationary frame, this wander from the (now moving) Lagrange point is depicted in Figure \ref{fig:greeksdeviationstationaryframe}. The deviation oscillates with a magnitude of $9.10 \times 10^{-2}$ AU, and a period of $148 \pm 4$ years. This is modulated with a faster oscillation component of $11.85 \pm 0.27$ years, equivalent to the orbital period of the asteroids.

@@

These much more significant errors are due to time dependence in the exact solution, as detailed in Section \ref{theory}. Energy can also be evaluated, and conserved, in this inertial frame; asteroid specific energy varies within only 0.113\% of the initial value, and with a similar periodicity to wander. Energy is also negative, confirming the asteroids are located within a bound orbit.

Animations produced to demonstrate the orbit in the stationary frame are included in Supplementary Material I-II. Animation I depicts the orbit as evaluated by the radau solver, while II depicts it with LSODA, demonstrating the drift present over time with non-stiff solvers.

\subsection{Wander Analysis} \label{perturbed}
The wander from the initial point was calculated for random perturbations with a maximum magnitude of 1\% of the displacement from the origin, and considered  separately perturbation components parallel and perpendicular to the position vector from the CoM (hereafter referred to as radial and tangential components respectively). By considering these perturbations across position space, it was clear that the wander is fully determined by the radial component, with no tangential dependence (as shown in Figure \ref{fig:wanderplots}).

@@

Figure \ref{fig:wander_rad} demonstrates a polynomial dependence on perturbation size in the radial direction, with a negligible constant term. While a quadratic has been fitted here, it was not possible to eliminate the possibility of higher order terms, as increasing perturbation size beyond 0.06 AU can lead to unstable orbits.

We may consider the wander resulting from perturbations in position and velocity space. Figure \ref{fig:position_mesh} clearly shows the dependence on radial position perturbations only, while Figure \ref{fig:velocity_mesh} shows a similar the dependence on tangential velocity perturbations.

@@

\begin{figure}[ht]
	\centering
	\begin{subfigure}{.23\textwidth}
		\centering
		% include first image
		\includegraphics[width=\linewidth]{Figures/radialp_orbits}  
		\caption{Radial}
		\label{fig:orbit_rad}
	\end{subfigure}
	\hfill %% useful if width of each figure is less the .5\textwidth
	\begin{subfigure}{.23\textwidth}
		\centering
		% include second image
		\includegraphics[width=\linewidth]{Figures/tangentialp_orbits}  
		\caption{Tangential}
		\label{fig:orbit_tan}
	\end{subfigure}
	\caption{Orbits from an initial perturbation of magnitude 0.01 AU from the origin, in radial and tangential directions, over 12 orbital periods. Note the 'tadpole' features resulting from the radial perturbation, and the significantly larger wander than from the tangential perturbation.}
	\label{fig:orbitplots}
\end{figure}

The "tadpole" orbit in Figure \ref{fig:orbit_rad} consists of two oscillating components, as described in Section \ref{lagrange}. The short period, measured at $11.85 \pm 0.61$ years, is in excellent agreement with  planetary orbital period as expected; meanwhile the long period is measured to be $148 \pm 5$ years, consistent with the analytical value of $144$ years.

For a narrow range of larger perturbations, stable horseshoe orbits encompassing both Lagrange points may also be observed as depicted in Figure \ref{fig:horseshoe}. This has a full period of $353$ years, in agreement with Taylor et al.'s numerical result of $358 \pm 7$ years \cite{Taylor1981}.

@@

\subsection{Perturbations in Z-direction} \label{3D}
Perturbations in the Z-direction are aligned with angular velocity vector for rotating frame, so experience no virtual forces in this direction, and oscillate under the influence of gravity alone. Considering small perturbations in the z direction (so that the distance from the CoM can be considered constant), the period of such oscillations tends to the orbital period of the rotating frame, under Newtonian gravity.

These are measured to be $11.85 \pm 0.48$ years, in strong agreement with the theoretical predictions.

@@

\begin{figure}
	\centering
	\includegraphics[width=0.8\linewidth]{Figures/wanderwithplanetmass_p5b}
	\caption{Variation of maximum wander with planetary mass, under a 0.01 AU radial perturbation and 0.01 AU/year tangential boost. Note the well defined quadratic trend up to $ M_{p} = 0.026 M_{s}$, and instability beyond this.}
	\label{fig:planetmass}
\end{figure}

A quadratic trend (with a large quadratic component) is initially observed, but wander deviates from this at approximately $ M_{p} = 0.026 M_{s}$, with a sharp increase indicating orbit instability. This occurs significantly before the theoretical prediction of $ M_{p} = 0.04 M_{s}$ \cite{Darwin1897}, however it is suggested this is due to instabilities in the integration solver.

\section{Conclusion}

@@

\onecolumn
\begin{appendices}
\section{Program Structure} \label{structure}
Global constants, along with derived values from these are given in an python module "\textit{constants}", from which relevant variable are imported in all other scripts.

Functions to evaluate these coupled differential equation systems are defined in module "\textit{orbits}", while additional functions to evaluate the wander during the orbit are implemented in "\textit{wander}". These modules are both imported into the scripts below for plotting and analysis, and given in the listings in Appendix \ref{Code}.

\textit{Wander from unperturbed initial conditions is calculated within "\textit{lagrange\_stability}", with ODE solver performance also evaluated here.
Orbits in the stationary frame are animated and saved within "\textit{stat\_frame\_video}".

2D perturbations are characterised in \textit{perturbation\_2D}, while perturbations in the z direction are considered in \textit{perturbation\_3D}. Variation in planetary mass, for results in section \ref{planet}, are given in \textit{planet\_mass\_variation}.}

Other code files are listed by Section below in Table \ref{Codefiles}:

>>>>

%\author{
	%\authorstyle{Christopher Gallagher}}
\author{\authorstyle{Christopher Gallagher} %Candidate 6952R}
	\institution{University of Cambridge}}
% Example of a one line author/institution relationship
%\author{\newauthor{John Marston} \newinstitution{Universidad Nacional Autónoma de México, Mexico City, Mexico}}

\date{\today} % Add a date here if you would like one to appear underneath the title block, use \today for the current date, leave empty for no date
\usepackage[english]{babel}
\usepackage{tabularx}
%\usepackage[backend=biber,doi=false]{biblatex}
\addbibresource{references.bib}
%\AtBeginBibliography{\small}

\usepackage{amsmath}
\usepackage{bm}
\usepackage{fixltx2e}% for \textsubscript

%----------------------------------------------------------------------------------------
\begin{document}

@@

\lettrineabstract{Trojan asteroids orbit the Sun in resonance with Jupiter at Lagrange points,
	and have their own stable orbits about these points. This can be treated as a restricted three-body problem, which is numerically evaluated in this report using a stiff Radau method to solve the system of coupled differential equations. Without perturbations, a maximal deviation of 4.68\pmb{$\times$}10\pmb{$^{-13}$} AU  from the Lagrange points was observed over 1000 orbits (11,852 years). The application of perturbations to these asteroids demonstrated the predicted existence of stable tadpole and horseshoe orbits, and suggested that wander from the Lagrange point depends quadratically on perturbation size. Such orbits were observed to be stable up to a planetary mass of 0.04 \pmb{$M_{\odot}$}, also in agreement with theoretical predictions. Word Count: XXXX}

@@

%Use https://leancrew.com/all-this/2016/08/lagrange-points-redux/ to create your own contour plot?
\section{Introduction}
The Jupiter Trojans, commonly known as the Trojan asteroids, are two large groups of asteroids that share the planet Jupiter's orbit around the Sun in a 1:1 orbit resonance. These two groups are called the Greeks and the Trojans, named after opposing sides in the mythological Trojan war, and lead/trail Jupiter respectively in its orbit. They correspond to Jupiter's two stable Lagrange points: L\subscript{4}, lying 60° ahead of the planet in its orbit, and L\subscript{5}, 60° behind, with asteroids distributed in two elongated, curved regions around these Lagrange points.

The first Jupiter Trojan, 588 Achilles, was discovered in 1906 by the German astronomer Max Wolf \cite{Nicholson1961}, and a total of 7642 Jupiter Trojans have been found as of February 2020 \cite{IAU2020}.

Research into Jupiter's Trojan asteroids continues, with the particular focus on their origins reliant on an understanding of their orbit stability \cite{DiSisto2019} \cite{Nesvorn2018}. This informs studies into their composition \cite{Brown2016}, as travel to these asteroids is considered for their potential in mineral mining \cite{Okada2017} \cite{Levison2016}.

The purpose of this report is to use numerical simulation techniques to investigate the stability of orbits about the Lagrange points, demonstrating that asteroids oscillate about these points with small perturbations and quantifying the absolute distance of the asteroids from the Lagrange point (the wander) during their orbits. The impact of variation in planetary/solar mass on asteroid orbit stability will also be considered.

@@

\section{Theoretical Background}

\subsection{Lagrange Points} \label{lagrange}
The asteroids exist at or near Lagrange points, defined in Lagrange's initial analysis of the three-body problem in 1772 \cite{Lagrange1772}, in which he demonstrated the existence of five equilibrium points for an object of negligible mass orbiting under the gravitational effect of two larger masses. Three of these equilibrium points, L\subscript{1}\textendash L\subscript{3} lie on the line joining the two masses, while each of the remaining two points, L\subscript{4} and L\subscript{5}, lie at the apex of an equilateral triangle with base equal to the separation of the two masses (see Figure \ref{fig:lagrangepoints}). Although these points are all potential maxima, stable motion is possible around L\subscript{4} and L\subscript{5} due to the Coriolis force \cite{Lissauer2014}.
%lissauer also says stability is provided that the most massive body has at least 25 times the mass of the secondary

\begin{figure}[h]
	\centering
	\includegraphics[width=\linewidth]{Figures/lagrange_points}
	\caption{The location of the five Lagrange equilibrium points in the circular-restricted three-body problem. The solar and planetary masses are denoted by the large and small filled circles, and the letters P, H, and T denote passing, horseshoe, and tadpole orbits respectively. Note that the two massive bodies form an equilateral triangle with each of the L\subscript{4} and L\subscript{5} points. Reproduced from Marzari et al. \cite{Marzari2002}}
	\label{fig:lagrangepoints}
\end{figure}

Figure \ref{fig:lagrangepoints} depicts orbits about L\subscript{4} and L\subscript{5} (known as tadpole orbits), as well as horseshoe orbits between Lagrange points, described by Murray et al. \cite{Murray1999}.

This report will focus on tadpole orbits, a well-documented feature of Trojan orbits \cite{Garfinkel1983} \cite{Dermott1981} and of the restricted three-body problem in general. Their distinctive shape results from a long-period motion about the equilibrium point combined with a short-period oscillation due to the Keplerian motion of the asteroid.

Szebehely et al. \cite{Szebehely1969} predict that this short period tends to the planetary period in the small planetary mass limit, while the long period is given by

\begin{equation}
T_{long} = T_{P} \sqrt{\frac{4}{27 \mu_{2}}}  
\end{equation}

@@

\subsection{Theoretical Model} \label{theory}
The three-body problem, where the dynamics of three interacting bodies are determined from their initial positions and velocities, has no analytical (closed-form) solution in the general case \cite{Barrow2008}.

@@

The system of differential equations determines the position and velocity of the asteroids, with two equations per spatial coordinate.
\begin{equation}
\frac{dr_{i}}{dt} = v_{i}, \frac{dv_{i}}{dt} = g_{i}, \quad i = x,y 
\end{equation}

@@

We may also consider a frame rotating at the same velocity as the massive bodies. As there is 1:1 orbital resonance between Jupiter and the asteroids, all three bodies are stationary in this frame. This significantly increases the accuracy of numerical simulations, as the exact solution is stationary with no explicit time dependence, rather than requiring an infinite power series \cite{Guglielmi2001}, \cite{LeVeque2007}.

@@

\subsection{Symmetry} \label{symmetry}
This problem contains a number of symmetries, which were employed to simplify the problem and reduce the computational load. The Trojan and Greek asteroids are in equivalent positions, so experience the same forces, and the system has rotational and inversion symmetry, so the choice of initial point and orbit direction is arbitrary. Therefore, only the Greeks, orbiting counter-clockwise with perturbations applied at t = 0, need be considered.

\subsection{Orbit Geometry}
As the three bodies considered here form an equilateral triangle in the initial equilibrium state, as depicted in Figure \ref{fig:geometry}, we can easily derive the polar coordinates of each body with respect to the centre of mass about which the bodies orbit.

\begin{figure} %place in text?
	\centering
	\resizebox{\columnwidth}{!}{%
		\begin{tikzpicture}
		\draw (0,0) -- (6,0) -- (3, {6*sin(60)} ) node[above]{Asteroids} node[midway,right]{$a$} -- cycle node[midway,left]{$a$};
		\draw  (1,0) arc (0:60:1cm)  ;
		%\node [label={[shift = {(10mm,5mm)}]{$\frac{\pi}{3}$}}] {};
		\node [label={[shift = {(13:11mm)}]{$\frac{\pi}{3}$}}] {};
		\node [label={[shift = {(-7mm,-7mm)}]{Sun}}] {};
		\node [label={[shift = {(67mm,-7mm)}]{Planet}}] {};

@@

		\end{tikzpicture}
	}	
	\caption{A geometric depiction of the three-body system, in the case where the planet has a mass equal to one-third of the sun, and asteroids are considered at the L\subscript{4} point. $ R_{s} $ and $ R_{p}$ denote the (fixed) radii from the centre of mass (the grey point) to the Sun and the planet respectively, while $ r_{a} $ denotes the radius of the asteroids.}
	\label{fig:geometry}
\end{figure}

@@

\begin{equation} \label{frequency}
\Omega = \sqrt{\frac{G (M_{s} + M_{p})}{a^{3}}}
\end{equation}

\section{Methodology}
This system of coupled first-order ordinary differential equations (ODEs) was solved using the scipy \textit{solve\_ivp} function. The time span was taken as 100 orbits (with 100 points sampled per orbit) unless otherwise stated; this corresponds to 1185 Earth years. Rescaled solar system units are used for mathematical ease, so distances are measured in astronomical units (AU), time in Earth years and mass in multiples of the solar mass, to prevent floating point overflows due to the magnitude of the quantities considered in SI units.

@@

\subsection{Integration Method}
Within \textit{solve\_ivp}, the default solver is RK45 (an explicit Runge-Kutta method of order 5(4) \cite{Dormand1980}) which gives a deviation in asteroid position (from the Lagrange point) in the order of $ 10^{-4}$ AU in the rotating frame over 50 orbits. This is larger than expected, suggesting that the system of equations may require an unreasonably small step-size for numerical stability in this method, even in regions where the solution curve is smooth \cite{Lambert1991}. Such systems are known as stiff, and solvers designed for these systems typically do more work per step, allowing them to take much larger steps, and have improved numerical stability compared to non-stiff solvers, such as RK45 \cite{Byrne1987}.

Therefore the stiff Radau solver (an implicit Runge-Kutta method from the Radau IIA family of order 5 \cite{Hairer2010}) is used for increased stability \cite{Frank1985}, and achieves a deviation in asteroid position in the order of $ 10^{-13}$ AU instead. This also ensures stability in the stationary frame, with deviations of 0.76\% in asteroid separation from Jupiter over $ 10^{3} $ years, compared to 53\% for the best non-stiff solvers.

@@

\subsection{Programme Structure}
Global constants, such as solar mass, and sun\textendash planet separation, along with derived values from these, such as orbital period and solar radius from the CoM, are given in an importable python module, "\textit{constants}".

Functions to evaluate these coupled differential equation systems are defined in the module "\textit{orbits}", while additional functions to evaluate the wander during the orbit (under different sampling routines) are implemented in "\textit{wander}". Complete code listings for these files are given in Appendix \ref{Code}. Further files then import these modules and produce the plots given in this report, fully detailed in Appendix \ref{structure}.

When varying planetary mass, I considered it preferable to avoid reconstructing all functions to take planetary mass as an argument, as this requires re-evaluating all initial derived constants in the "\textit{orbits}" module. Therefore, I iterate over alternative planetary masses, re-defining constant values in each instance, and directly import the required functions to compute the orbit.

\subsection{Performance} \label{Performance}
Sampling 100 points per orbit for 100 orbital periods takes a mean time of $16.97 \pm 34$ \si{\milli\second}, with sub-linear scaling for sampling rate and approximate linear scaling with orbit number up to the array memory limit, achieved through the optimised integration routines within \textit{solve\_ivp}. Computing the wander over position/velocity space with larger perturbations was, however, more time consuming, averaging $2.04 \pm 0.16$ \si{\second} per point with the same parameters.

The main computational load was within the \textit{solve\_ivp} function, which is already optimised well beyond the capabilities of the author. However, assumptions made in Section \ref{symmetry}, such as considering only one group of asteroids, reduces this time somewhat, and vectorisation of other aspects of the \textit{orbit} functions reduced the running time of the whole module. These approaches were not taken universally however, due to the size of the arrays required, and the dominant effect of the ODE solver on the running time; instead, the orbit number was reduced when only a comparison of the maximum wander was required.

\subsection{Fourier Analysis} \label{Fourier}
For periodic oscillations and orbits, the Fast Fourier Transform (FFT) implemented in \textit{scipy} is used to obtain the period, with \textit{find\_peaks} in \textit{scipy.signal} identifying the exact values. The errors were estimated using the associated peak widths of each frequency component.

\section{Results}
\subsection{Unperturbed Stability of Lagrange Points} \label{unperturbed}
Without perturbations applied, the Greeks' orbit has a maximum deviation from L\subscript{4} of $4.68 \times 10^{-13}$ AU over 100 orbits (1185 years). This value is unchanged if 1000 orbits are considered instead, confirming the stability of this Lagrange point.

In the stationary frame, this wander from the (now moving) Lagrange point is depicted in Figure \ref{fig:greeksdeviationstationaryframe}. The wander oscillates with a magnitude of $9.10 \times 10^{-2}$ AU, over a period of $148 \pm 4$ years. This is modulated with a faster oscillation component of $11.85 \pm 0.27$ years, equivalent to the orbital period of the asteroids.

@@

These much more significant deviations are due to time-dependence in the exact solution, as detailed in Section \ref{theory}. Energy can also be evaluated, and conserved, in this inertial frame; asteroid specific energy varies within only 0.113\% of the initial value, and with a similar periodicity to wander. Energy is also negative, confirming the asteroids are located within a bound orbit.

Animations produced to demonstrate the orbit in the stationary frame are included in Supplementary Material I-II. Animation I depicts the orbit as evaluated by the Radau solver, while II depicts it with LSODA, demonstrating the drift present over time with non-stiff solvers.

\subsection{Wander Analysis} \label{perturbed}
Wander from the initial point was calculated for random perturbations with a maximum magnitude of 1\% of the displacement from the origin. The perturbation components parallel and perpendicular to the position vector from the CoM (hereafter referred to as radial and tangential components respectively) were considered separately. By considering these perturbations across position space, it was clear that the wander is fully determined by the radial component, with no tangential dependence (as shown in Figure \ref{fig:wanderplots}).

@@

Figure \ref{fig:wander_rad} demonstrates a polynomial dependence on perturbation size in the radial direction, with a negligible constant term. While a quadratic has been fitted here, the possibility of higher order terms could not be eliminated, as increasing perturbation size beyond 0.06 AU can lead to unstable orbits.

We may consider the wander resulting from perturbations in position and velocity space. Figure \ref{fig:position_mesh} clearly shows dependence on radial position perturbations only, while Figure \ref{fig:velocity_mesh} shows a similar dependence on tangential velocity perturbations.

@@

\begin{figure}[ht]
	\centering
	\begin{subfigure}{.23\textwidth}
		\centering
		% include first image
		\includegraphics[width=\linewidth]{Figures/radialp_orbits}  
		\caption{Radial}
		\label{fig:orbit_rad}
	\end{subfigure}
	\hfill %% useful if width of each figure is less the .5\textwidth
	\begin{subfigure}{.23\textwidth}
		\centering
		% include second image
		\includegraphics[width=\linewidth]{Figures/tangentialp_orbits}  
		\caption{Tangential}
		\label{fig:orbit_tan}
	\end{subfigure}
	\caption{Orbits from an initial perturbation of magnitude 0.01 AU from the origin, in radial and tangential directions, over 12 orbital periods. Note the tadpole features resulting from the radial perturbation, and the significantly larger wander than from the tangential perturbation.}
	\label{fig:orbitplots}
\end{figure}

The tadpole orbit in Figure \ref{fig:orbit_rad} consists of two oscillating components, as described in Section \ref{lagrange}. The short period, measured at $11.85 \pm 0.61$ years, is in excellent agreement with  planetary orbital period as expected; meanwhile the long period is measured to be $148 \pm 5$ years, consistent with the analytical value of $144$ years.

For a narrow range of larger perturbations, stable horseshoe orbits encompassing both Lagrange points may also be observed, as depicted in Figure \ref{fig:horseshoe}. This has a full period of $353$ years, in agreement with Taylor et al.'s numerical result of $358 \pm 7$ years \cite{Taylor1981}.

@@

\subsection{Perturbations in Z direction} \label{3D}
Perturbations in the Z direction are aligned with the angular velocity vector for the rotating frame, so experience no virtual forces in this direction, and oscillate under the influence of gravity alone. Considering small perturbations in the Z direction (so that the distance from the CoM can be considered constant), the period of such oscillations tend to the orbital period of the rotating frame under Newtonian gravity, as detailed in Appendix \ref{zperiod}. These oscillations are observed to be approximately sinusoidal, with a measured period of $11.85 \pm 0.48$ years, in strong agreement with the theoretical predictions.

\begin{figure}
	\centering
	\includegraphics[width=0.8\linewidth]{Figures/z_perturbations5c}
	\caption{Variation of wander with increasing Z perturbations. A strong linear component is observed initially in the un-perturbed case, but then tends towards the perturbed quadratic case, while the linear offsets result from the initial perturbation in the XY plane.}
	\label{fig:zwander}
\end{figure}

The overall wander, however, is observed to follow Figure \ref{fig:zwander}.
It is suggested that, while the maximum deviation from the Lagrange point in the Z direction is simply twice the initial perturbation, wander in the XY plane increases quadratically. This can be initiated from a radial displacement due to the reduced gravitational force  under the Z perturbation, demonstrated in Figure \ref{fig:zorbit} or, more significantly, from an initial perturbation in the XY plane. However, the oscillation of gravitational forces due to perturbations in the Z direction make this problem significantly more complicated and are a possible cause of the observed deviations from the quadratic relations.

\begin{figure}
	\centering
	\includegraphics[width=0.8\linewidth]{Figures/z_perturb_orbit}
	\caption{Orbit of Greek asteroids following a 0.2 AU perturbation in the positive Z direction, over 12 orbital periods. Note the initial outwards radial displacement despite the lack of perturbation in the XY plane, and subsequent tadpole-like orbit.}
	\label{fig:zorbit}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.8\linewidth]{Figures/wanderwithplanetmass_p6}
	\caption{Variation of maximum wander with planetary mass, under a 0.01 AU radial perturbation and 0.01 AU/year tangential boost. Note the orbit stability up to $ M_{p} = 0.04 M_{s}$, and instability beyond this.}
	\label{fig:planetmass}
\end{figure}

An approximate quadratic trend is initially observed in the small mass region, with minimum wander occurring at $ M_{p} = 0.026 M_{s}$. However, the variation of wander below $ M_{p} = 0.04 M_{s}$ is minimal, while orbits are unstable beyond this point, in agreement with theoretical predictions by Darwin \cite{Darwin1897}.

\section{Conclusion}
This report demonstrates the stability of Lagrange points L\subscript{4} and L\subscript{5}, with a maximum deviation from the Lagrange points of $4.68 \times 10^{-13}$ AU over 1000 orbits (11,852 years), in a frame rotating with the asteroids and without perturbations applied. Stability under perturbations in both the XY plane and in the Z direction is also demonstrated, with a finite wander from these points independent of orbit duration. It was possible to further quantify this wander, suggesting a strong quadratic dependence on radial perturbation magnitude and independence of tangential perturbations, which did not grow beyond the initial perturbation size.
It was also possible to replicate both tadpole and horseshoe orbits from radial perturbations, and frequency components were obtained for these orbits in agreement with previous literature. It was also found that orbits become unstable when the planetary\textendash solar mass ratio is greater than 0.04, in agreement with theoretical predictions, and there is little mass dependence below this point.

Further work is required to evaluate coefficient values for polynomial relationships observed, and to demonstrate a theoretical basis for such dependencies.
Further research would also be beneficial on the categorisation of different orbit types, and examination of initial conditions corresponding to formation mechanisms of these orbits.

@@

\onecolumn
\begin{appendices}
\section{Z Perturbation Period} \label{zperiod}
Figure \ref{fig:geometryz} depicts a Z perturbation out of the XY plane, from the L\subscript{4} point.
\begin{figure} [h]
	\centering
	\resizebox{0.5\columnwidth}{!}{%
		\begin{tikzpicture}
		\draw (0,0) -- (6,0)  node[midway, below]{$a$} node[right]{L\subscript{4}}; 
		\draw [arrows=->] (6,0) --  (6, 1.95 ) node[above]{Asteroids} node[midway,right]{$z_{0}$};
		\draw [dashed] (6,2) -- (0,0) node[midway, above]{$R$};
		\node [label={[shift = {(-7mm,-7mm)}]{Sun}}] {};
		\fill[black] (0,0) circle (3.5mm);% node[anchor=north west] {Sun};
		\fill[black] (6,2) circle (0.5mm);
		\fill[gray] (6,0) circle (0.5mm);
